---
title: "R plot 2d pixel"
author: "Jun-Liszt Li"
date: 2019-05-01T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown", "2d density", "ggplot2"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# 2d density plot with ggplot2

A 2d density plot is useful to study the relationship between 2 numeric variables if you have a huge number of points.

To avoid overlapping (as in the scatterplot beside), it divides the plot area in a multitude of small fragment and represents the number of points in this fragment.

There are several types of 2d density plots. Each has its proper ggplot2 function. This post describes all of them.

You can embed an R code chunk like this:

```{r 2d}
# Library
library(tidyverse)
 
# Data
a <- data.frame( x=rnorm(20000, 10, 1.9), y=rnorm(20000, 10, 1.2) )
b <- data.frame( x=rnorm(20000, 14.5, 1.9), y=rnorm(20000, 14.5, 1.9) )
c <- data.frame( x=rnorm(20000, 9.5, 1.9), y=rnorm(20000, 15.5, 1.9) )
data <- rbind(a,b,c)
 
 
# Basic scatterplot
ggplot(data, aes(x=x, y=y) ) +
  geom_point()
library(ggplot2)
# Show the contour only
ggplot(data, aes(x=x, y=y) ) +
  geom_density_2d()
 
# Show the area only
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon")
 
# Area + contour
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white")
 
# Using raster
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
```

# Customize the color palette

Whatever you use a 2d histogram, a hexbin chart or a 2d distribution, you can and should custom the colour of your chart. Here is a suggestion using the scale_fill_distiller() function. You can see other methods in the ggplot2 section of the gallery.

```{r color}
# Call the palette with a number
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette=4, direction=-1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
 
# The direction argument allows to reverse the palette
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette=4, direction=1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
 
# You can also call the palette using a name.
ggplot(data, aes(x=x, y=y) ) +
  stat_density_2d(aes(fill = ..density..), geom = "raster", contour = FALSE) +
  scale_fill_distiller(palette= "Spectral", direction=1) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    legend.position='none'
  )
```


# Reference\

1. https://r-graph-gallery.com/2d-density-plot-with-ggplot2.html#hex
2. https://journal.r-project.org/archive/2013-1/kahle-wickham.pdf

```{r}

```

